---
title: "Untitled"
author: "Pavitra Desai"
date: "23/10/2021"
output:
  html_document:
    keep_md: yes
---
### Introduction 
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.The goal of this project is to predict the manner in which they did the exercise.  
### Requirements
```{r}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
```
### Downloading data
```{r}
trainurl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testurl  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainurl))
testing <- read.csv(url(testurl))
dim(training)
dim(testing)
```
### Cleaning data
```{r}
nzv <- nearZeroVar(training)
training <- training[,-nzv]
testing <- testing[,-nzv]
dim(training)
dim(testing)
isNA <- sapply(training, function(x) mean(is.na(x))) 
training<- training[,isNA == FALSE]
testing <- testing[,isNA == FALSE]
dim(training)
dim(testing)
training <- training[,8:59]
testing <- testing[,8:59]
dim(training)
dim(testing)
sub <- createDataPartition(training$classe, p=0.6, list=FALSE)
training <- training[sub,]
testing <- training[-sub,]
dim(training)
dim(testing)
```
### Decision Tree Model
```{r}
training$classe=factor(training$classe)
DT_modfit <- train(classe ~ ., data = training, method="rpart")
DT_prediction <- predict(DT_modfit, testing)
confusionMatrix(table(DT_prediction, testing$classe))
rpart.plot(DT_modfit$finalModel, roundint=FALSE)
```
### Random Forest Model
```{r}
training$classe=factor(training$classe)
modFitRF <- randomForest(classe ~ ., data=training, method="class")
# Perform prediction
predictRF <- predict(modFitRF, testing, type = "class")
confusionMatrix(table(predictRF, testing$classe))
```
### Conclusion
The confusion matrices show, that the Random Forest algorithm performens better than decision trees. The accuracy for the Random Forest model was ~1 compared to 0.497 for Decision Tree model. The random Forest model is choosen.
